Designing a high-performance boundary element library with OpenCL and Numba

### 2021-04-06 opencl.tex
Further work

>>>>

\subsection{Assembling the singular part of integral operators}

The assembly of the singular part of an integral operator works a bit differently. Remember that the singular part consists of triangle parts which are adjacent to each other or identical. There are $\mathcal{O}(N)$ of such pairs. We are using fully numerical quadrature rules for these integrals that are based on subdividing the four-dimensional integration domain and using transformation techniques to remove the singularities. This gives highly accurate evaluations of these integration pairs but requires a large number (typically over $1000$) quadrature points per triangle pair. Hence, we create one workgroup for each triangle pair and inside this workgroup have a number of work-items that are evaluating the quadrature rules and then at the end sum up the result together. Depending on how two triangles are in relationship to each other different types of singular quadrature rules are needed. We solve this pre-loading all possible quadrature rules onto the device and also store for each triangle pair an index into the required quadrature rule so that the kernel function can select the correct rule to evaluate. For the singular quadrature rules we did not implement separate SIMD optimized kernels as the proposed implementation is already highly efficient and requires only a fraction of the computational time of the regular quadrature rules described above. At the end, the singular integral values are either summed into the overall result matrix, or if desired by the user, stored as separate sparse matrix.

\section{Numba assembly of integral operators}

The Numba assembly of integral operators is structured differently from the OpenCL assembly. In OpenCL we create a large number of work-items that are executed in parallel on the compute device by the runtime system with the details of the parallel execution left to the device driver. For Numba we use classical loop based parallelism. Each loop iteration is the assembly of one test element with all trial elements. Within each loop we partially optimized for SIMD instruction, though this depends on the autovectorization of the just-in-time compiler and cannot be explicitly targeted in Numba. Again, we separate regular and singular interactions. In the next section we give some more details on SIMD optimization in Numba at the example of the Laplace kernel.

\section{SIMD optimized evaluation of the Laplace kernel with Numba}
